%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% abstrac-en.tex
%% NOVA thesis document file
%%
%% Abstract in English([^%]*)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE abstrac-en.tex}

Regardless of the language in which the dissertation is written, a summary is required in the same language as the main text and another summary in another language. It is assumed that the two languages in question are Portuguese and English.

The abstracts should appear first in the language of the main text and then in the other language. For example, if the dissertation is written in Portuguese the abstract in Portuguese will appear first, then the abstract in English, followed by the main text in Portuguese. If the dissertation is written in English, the abstract in English will appear first, then the abstract in Portuguese, followed by the main text in English. 

In the \LaTeX\ version, the NOVAthesis template will automatically order the two abstracts taking into account the language of the main text. You may change this behaviour by adding
\begin{verbatim}
    \abstractorder(<MAIN_LANG>):={<LANG_1>,...,<LANG_N>}
\end{verbatim}
\noindent to the customization area in the document preamble, e.g.,
\begin{verbatim}
    \abstractorder(de):={de,en,it}
\end{verbatim}

The abstracts should not exceed one page and, in a generic way, should answer the following questions (it is essential to adapt to the usual practices of your scientific area):

\begin{enumerate}
  \item What is the problem?
  \item Why is this problem interesting/challenging?
  \item What is the proposed approach/solution?
  \item What results (implications/consequences) from the solution?
\end{enumerate}

Feature selection methods for cancer classification
are aimed to overcome the high dimensionality of the biomedical
data which is a challenging task. Most of the feature selection
methods based on DNA methylation are time consuming during
testing phase to identify the best pertinent features subset that
are relevant to accurate prediction. However, the hybridization
between feature selection and extraction methods will bring a
method that is far fast than only feature selection method. This
paper proposes a framework based on both novel feature
selection methods that employ statistical variation, standard
deviation and entropy, along with extraction methods to predict
cancer using three new features, namely, Hypomethylation,
Midmethylation and Hypermethylation. These new features
represent the average methylation density of the corresponding
three regions. The three features are extracted from the selected
features based on the analysis of the methylation behavior. The
effectiveness of the proposed framework is evaluated by the
breast cancer classification accuracy. The results give 98.85%
accuracy using only three features out of 485,577 features. This
result proves the capability of the proposed approach for breast
cancer diagnosis and confirms that feature selection and
extraction methods are critical for practical implementation. 



% Palavras-chave do resumo em Inglês
\begin{keywords}
Keyword 1, Keyword 2, Keyword 3, …
\end{keywords} 
